---
title: "Machine Efficiency Example"
author: "Matthew Kuperus Heun"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{machine_effiency_example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  # Set several aspects of graph chunks
  fig.retina=4,
  fig.align='center', 
  fig.height=4,
  fig.width=6
)
```

```{r setup}
library(CLPFUDecompositionDatabase)
library(PFUSetup)
library(purrr)
```


## Introduction

This vignette provides an introduction to using the machine efficiency data
from the CL-PFU database and the CLPFUDecomposition calculation pipeline.


## Create the pinboard

The first step is to establish a "pinboard" 
for the [pins](https://pins.rstudio.com) package.
Someone from the MR-PFU Database team will have shared the "PipelineReleases"
folder if you have access to IEA data. 
The `PFUSetup` package is helpful to find the correct folder.
Install the `PFUSetup` package with the following code.

```{r, eval=FALSE}
devtools::install_github("EnergyEconomyDecoupling/PFUSetup")
```

Create the pinboard from the correct Dropbox folder.
Your `pinboard_folder` might be different from this example, 
due to a different location on your computer.

```{r}
pinboard_folder <- PFUSetup::get_abs_paths()[["pipeline_releases_folder"]]
pinboard_folder
pinboard <- pins::board_folder(pinboard_folder, versioned = TRUE)
pinboard
```


## Read machine efficiency data

Machine efficiency data is stored in the "eta_i" pin,
so named after the function in the `Recca` package that
performs the calculations [Recca::calc_eta_i()].
The version below contains efficiencies 
for the USA for the years 1960--2000.

```{r}
eta_i_data <- pins::pin_read(board = pinboard,
                             name = "eta_i",
                             version = "20230904T183344Z-d3924") |> 
  PFUPipelineTools::tar_ungroup()
names(eta_i_data)
```

The `eta_i` column contains machine efficiencies
as column vectors in each row. 
To unpack the efficiencies out of the vectors,
use the `matsindf` package.

```{r}
unpacked_eta_i_data <- eta_i_data |> 
  dplyr::filter(IEAMW == "IEA", Last.stage == "Final") |> 
  # Delete columns containing original PSUT matrices (if present)
  dplyr::mutate(
    R = NULL, U = NULL, U_feed = NULL, U_EIOU = NULL, r_EIOU = NULL,
    V = NULL, Y = NULL, S_units = NULL
  ) |> 
  # Put the data in the correct format for expanding the vectors
  tidyr::pivot_longer(cols = "eta_i", names_to = "matnames", values_to = "matvals") |> 
  # Expand the vectors
  matsindf::expand_to_tidy() |> 
  # Clean up
  dplyr::rename(
    machine = rownames, 
    eta_i = matvals
  ) |> 
  dplyr::mutate(matnames = NULL, colnames = NULL)
unpacked_eta_i_data
```


## Visualize the data

To create a graph for each machine, 
we can nest the data
to create an `eta_i` column that contains 
a `tibble` with year and efficiency columns.

```{r}
nested_eta_i_data <- unpacked_eta_i_data |> 
  tidyr::nest(.by = tidyr::all_of(c("Country", "Method", "Energy.type", "Last.stage", "IEAMW", "machine")), .key = "year_eta_i")
nested_eta_i_data
```

Finally, we can create a column of graphs.

```{r}
create_plot <- function(tibble_data, country_name, energy_type, machine_name) {
  ggplot2::ggplot() +
    ggplot2::geom_hline(yintercept = 0, linewidth = 0.1) +
    ggplot2::geom_hline(yintercept = 1, linewidth = 0.1) +
    ggplot2::geom_text(ggplot2::aes(x = 1990, y = -0.1, 
                                    label = paste(country_name, 
                                                   energy_type, 
                                                   machine_name, sep = "; "))) +
    ggplot2::geom_line(data = tibble_data,
                       mapping = ggplot2::aes(x = Year, y = eta_i)) +
    ggplot2::labs(x = NULL, y = expression(eta[i])) + 
    ggplot2::ylim(c(-0.5, 1.5)) +
    MKHthemes::xy_theme()
}
eta_i_graphs <- nested_eta_i_data |> 
  dplyr::mutate(
    # plots = purrr::map(.x = year_eta_i, .f = create_plot)
    plots = purrr::pmap(.l = list(tibble_data = year_eta_i, 
                                  country_name = Country, 
                                  energy_type = Energy.type,
                                  machine_name = machine), .f = create_plot)
  )
```

We can look at each graph.

```{r}
eta_i_graphs$plots[[1]]
eta_i_graphs$plots[[10]]
```
